Debugging by \textit{lastChange}

### 2010-11-02 trunk/papers/pldi2011/sigplanconf-template.tex
fixes for claude's comments

We have presented the \textit{lastChange} algorithm and described our
prototype implementation. Now we want to convince you that the
algorithm has practical value. Ultimately, practical value can only be
proven by developers in the field using the technique. This paper is
one step to convince implementers to put \textit{lastChange} in the
field.  Thus our goal here is a persuasive argument for potential implementors.

@@

Of course breakpoints are not feasible in these cases and logging
becomes unwieldy. Since automatic logging solutions are highly tuned,
their overhead in this case will likely be much less than
\textit{lastChange}. On the other hand, \textit{lastChange} integrates
with an interactive debugger and detecting that we are in a high
overhead loop is simply a matter of checking our internal counter. In
such unusual cases we may simply offer the developer the option of
studying the loop code then omitting it from \textit{lastChange}
operations. As a practical result, this is no different from similar
issues developers face with any debugger today: occasionally a
debugger causes too much overhead to be useful for debugging. On
balance we think \textit{lastChange} will rarely encounter these kinds
of issues.

@@

\textit{lastChange} functionality supports obtaining information about
the execution state logically earlier in the control flow. This
support resembles a mixture of replay-based and logging-based
debugging. Replay-based approaches capture limited data during
execution and replay the buggy execution to reach past points. In
contrast, logging-based approaches collect enough data during
execution to relieve developer from re-execution. Replay-based
approaches impose much less runtime overhead (about two orders of
magnitudes) comparing to logging-based approaches. However, developer
has to re-execute the buggy execution several
times. \textit{lastChange} functionality collects data on re-execution
but this data is limited to the current queries of developer.

>>>>

We have presented the \textit{lastChange} algorithm and described our
prototype implementation. Our goal is practical improvements in debugging. To achieve our goal we need practical JavaScript engines
to add new debug primitives so that developers in the field can use our new technique. This paper is
one step to convince implementers to enable \textit{lastChange}.
Thus we summarize here our arguments that \textit{lastChange} should be supported.

@@

Of course breakpoints are not feasible in these cases and logging
becomes unwieldy. Since automatic logging solutions are highly tuned,
their overhead in this case will likely be much less than
\textit{lastChange}. On the other hand, \textit{lastChange} integrates
with an interactive debugger and detecting that we are in a high
overhead loop is simply a matter of checking our internal counter. In
such unusual cases we may simply offer the developer the option of
studying the loop code then omitting it from \textit{lastChange}
operations. Developers face this issue with any debugger today: occasionally a
debugger causes too much overhead to be useful for debugging. On
balance we think \textit{lastChange} will rarely encounter these kinds
of issues.

@@

The algorithm we presented in this paper provides a way to obtain information about
the execution state logically earlier in the control flow. This
support resembles a mixture of replay-based and logging-based
debugging. Replay-based approaches capture limited data during
execution and replay the buggy execution to reach past points. In
contrast, logging-based approaches collect enough data during
execution to relieve developer from re-execution. Replay-based
approaches impose much less runtime overhead (about two orders of
magnitudes) comparing to logging-based approaches. However, developer
has to re-execute the buggy execution several
times. \textit{lastChange} functionality collects data on re-execution
but this data is limited to the current queries of developer.
